\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{titlesec}

% Compress section spacing to ensure it fits on one page
\titlespacing*{\section}{0pt}{1.2ex plus 1ex minus .2ex}{0.5ex plus .2ex}

% Remove paragraph indentation and add space between paragraphs for a cleaner block look
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5em}

\begin{document}

\begin{center}
    \Large \textbf{Drift-Aware Trajectory Control for Localized Diffusion Editing} \\[1\baselineskip]
    \normalsize Ali Özhavala, Edip Emre Dönger, Yiğit Kemal Can, Yüksel Barkın Baydar
\end{center}



\section{Introduction and Motivation}
Text-guided image editing with Latent Diffusion Models (LDMs) \cite{rombach2022high} enables high-quality localized modifications. Recent methods preserve non-edited regions via cross-attention manipulation \cite{hertz2022prompt}, mask guidance \cite{couairon2023diffedit}, or intermediate feature injection \cite{tumanyan2023plug, cao2023masactrl}. 

Despite these advances, \textit{background drift} often occurs during later denoising stages because fixed heuristic injection schedules (e.g., constant or early-only) cannot adapt as small deviations accumulate into unintended shifts. We propose reframing feature injection as a \textit{feedback regulation problem}, dynamically adjusting preservation strength based on a measured background drift signal at each timestep.

\section{Proposed Method and Contribution}
We introduce \textbf{Drift-Aware Adaptive Injection} for Stable Diffusion v1.5, operating entirely during inference without retraining.

\textbf{Baseline (Static Injection):} Following \cite{tumanyan2023plug}, we extract spatial feature maps from the UNet decoder during a reference and edited denoising trajectory. A softened spatial mask blends these features, injecting reference features into the background to preserve it. Baselines include constant, early-only, and linear decay schedules.

\textbf{Proposed (Adaptive Injection):} Instead of a fixed schedule, we measure background drift at each timestep by computing the divergence between intermediate feature representations of the reference and edited runs in the non-edited region. 
\begin{itemize}[noitemsep, topsep=2pt, leftmargin=*]
    \item \textbf{High Drift:} If features diverge, we dynamically increase preservation strength.
    \item \textbf{Stable Background:} If features align, we reduce intervention to avoid suppressing the edit.
\end{itemize}
We apply temporal smoothing to the drift estimate to prevent oscillatory injection. 

\section{Implementation and Evaluation Plan}
\textbf{Data \& Implementation:} We will curate an evaluation set of $\sim$40 Stable Diffusion-generated base images, paired with target prompts and spatial masks covering object replacements, attribute tweaks, and style changes. Using PyTorch, we will extract and blend paired UNet features with a shared noise seed. 

\textbf{Evaluation Metrics:} Background preservation will be evaluated using \textbf{LPIPS} and \textbf{SSIM} strictly outside the edit mask. Edit fidelity will be measured via \textbf{CLIP similarity} between the edited region crop and the target prompt. 

\textbf{Expected Outcomes:} We hypothesize that static injection degrades background structure late in generation by failing to respond to emerging drift. Our drift-aware approach will better stabilize the background by intervening precisely when deviations arise, maintaining competitive edit fidelity.

\vspace{1em}

\renewcommand{\refname}{\normalsize References}
\begin{thebibliography}{5}
\small
\bibitem{rombach2022high} 
Rombach, R., Blattmann, A., Lorenz, D., Esser, P., \& Ommer, B. (2022). High-Resolution Image Synthesis with Latent Diffusion Models. \textit{Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}.

\bibitem{hertz2022prompt} 
Hertz, A., Mokady, R., Tenenbaum, J., Aberman, K., Pritch, Y., \& Cohen-Or, D. (2022). Prompt-to-Prompt Image Editing with Cross-Attention Control. \textit{International Conference on Learning Representations (ICLR)}.

\bibitem{couairon2023diffedit} 
Couairon, G., Verbeek, J., Schwenk, H., \& Cord, M. (2023). DiffEdit: Diffusion-based Semantic Image Editing with Mask Guidance. \textit{International Conference on Learning Representations (ICLR)}.

\bibitem{tumanyan2023plug} 
Tumanyan, N., Geyer, M., Bagon, S., \& Dekel, T. (2023). Plug-and-Play Diffusion Features for Text-Driven Image-to-Image Translation. \textit{Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}.

\bibitem{cao2023masactrl} 
Cao, M., Wang, X., Qi, Z. X., Shan, Y., Qie, X., \& Hu, Z. (2023). MasaCtrl: Tuning-Free Mutual Self-Attention Control for Consistent Image Synthesis and Editing. \textit{Proceedings of the IEEE/CVF International Conference on Computer Vision (ICCV)}.
\end{thebibliography}

\end{document}